---
title: "Gene Summaries from RefSeq Database"
author: "Zuguang Gu (z.gu@dkfz.de)"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gene Summaries from RefSeq Database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package provides long description of genes collected from [the RefSeq database](https://ftp.ncbi.nih.gov/refseq/).
The text in "COMMENT" section started with "Summary:" is extracted as the description of the gene, e.g. in the following example:

```
LOCUS       NM_012363                936 bp    mRNA    linear   PRI 12-FEB-2021
DEFINITION  Homo sapiens olfactory receptor family 1 subfamily N member 1
            (OR1N1), mRNA.
ACCESSION   NM_012363 XM_071152
VERSION     NM_012363.1
KEYWORDS    RefSeq; MANE Select.
SOURCE      Homo sapiens (human)
  ORGANISM  Homo sapiens
            Eukaryota; Metazoa; Chordata; Craniata; Vertebrata; Euteleostomi;
            Mammalia; Eutheria; Euarchontoglires; Primates; Haplorrhini;
            Catarrhini; Hominidae; Homo.
REFERENCE   1  (bases 1 to 936)
  AUTHORS   Malnic B, Godfrey PA and Buck LB.
  TITLE     The human olfactory receptor gene family
  JOURNAL   Proc Natl Acad Sci U S A 101 (8), 2584-2589 (2004)
   PUBMED   14983052
  REMARK    Erratum:[Proc Natl Acad Sci U S A. 2004 May 4;101(18):7205]
REFERENCE   2  (bases 1 to 936)
  AUTHORS   Fuchs T, Malecova B, Linhart C, Sharan R, Khen M, Herwig R,
            Shmulevich D, Elkon R, Steinfath M, O'Brien JK, Radelof U, Lehrach
            H, Lancet D and Shamir R.
  TITLE     DEFOG: a practical scheme for deciphering families of genes
  JOURNAL   Genomics 80 (3), 295-302 (2002)
   PUBMED   12213199
REFERENCE   3  (bases 1 to 936)
  AUTHORS   Rouquier S, Taviaux S, Trask BJ, Brand-Arpon V, van den Engh G,
            Demaille J and Giorgi D.
  TITLE     Distribution of olfactory receptor genes in the human genome
  JOURNAL   Nat Genet 18 (3), 243-250 (1998)
   PUBMED   9500546
  REMARK    Erratum:[Nat Genet 1998 May;19(1):102]
COMMENT     REVIEWED REFSEQ: This record has been curated by NCBI staff. The
            reference sequence was derived from AL359636.17.
            On Apr 5, 2004 this sequence version replaced XM_071152.1.
            
            Summary: Olfactory receptors interact with odorant molecules in the
            nose, to initiate a neuronal response that triggers the perception
            of a smell. The olfactory receptor proteins are members of a large
            family of G-protein-coupled receptors (GPCR) arising from single
            coding-exon genes. Olfactory receptors share a 7-transmembrane
            domain structure with many neurotransmitter and hormone receptors
            and are responsible for the recognition and G protein-mediated
            transduction of odorant signals. The olfactory receptor gene family
            is the largest in the genome. The nomenclature assigned to the
            olfactory receptor genes and proteins for this organism is
            independent of other organisms. [provided by RefSeq, Jul 2008].
            
            ##RefSeq-Attributes-START##
            MANE Ensembl match     :: ENST00000304880.2/ ENSP00000306974.2
            RefSeq Select criteria :: based on single protein-coding transcript
            ##RefSeq-Attributes-END##
```

Function `loadGeneSummary()` extracts the gene summary table. Specifying the
`organism` argument with the full name or the corresponding taxon ID returns a
table of genes and their summaries:

```{r}
library(GeneSummary)
tb = loadGeneSummary(organism = 9606)
# # or use the full organism name
# tb = loadGeneSummary(organism = "Homo sapiens")
dim(tb)
head(tb)
```

Setting `organism` to `NULL` returns a table of all organisms.


```{r}
tb = loadGeneSummary(organism = NULL)
sort(table(tb$Organism))
sort(table(tb$Review_status))
```

A specific status can be set via argument `status`, e.g. only to `"reviewed"`:

```{r}
tb = loadGeneSummary(organism = NULL, status = "reviewed")
sort(table(tb$Review_status))
```

```{r}
sessionInfo()
```
